---
title: "Uber Statistics"
author: "Dhaval Sawlani"
date: "February 7, 2017"
output: html_document
---

```{r The Setup, cache=TRUE}
knitr::opts_chunk$set(warning = FALSE, fig.width = 9, fig.height = 6)
library(dplyr)
library(lubridate)
library(DT)
library(ggplot2)
library(RColorBrewer)
```

###Loading the Uber Data for the Year 2015

```{r, cache=TRUE}
uber_2015 <- read.csv("uber-raw-data-janjune-15.csv")
```

###Loading the Data for the Year 2014

```{r, cache=TRUE}
uber_apr <- read.csv("uber-raw-data-apr14.csv")
uber_may <- read.csv("uber-raw-data-may14.csv")
uber_jun <- read.csv("uber-raw-data-jun14.csv")
uber_jul <- read.csv("uber-raw-data-jul14.csv")
uber_aug <- read.csv("uber-raw-data-aug14.csv")
uber_sep <- read.csv("uber-raw-data-sep14.csv")
```

###Binding all the Data into one

```{r, cache=TRUE}
uber_2014 <- rbind(uber_apr, uber_may, uber_jun, uber_jul, uber_aug, uber_sep)
uber_2014$Date.Time <- mdy_hms(uber_2014$Date.Time)
uber_2014$Day <- factor(day(uber_2014$Date.Time))
uber_2014$Month <- factor(month(uber_2014$Date.Time, label = TRUE))
uber_2014$Year <- factor(year(uber_2014$Date.Time))
uber_2014$Weekday <- factor(wday(uber_2014$Date.Time, label = TRUE))
uber_2014$Hour <- factor(hour(uber_2014$Date.Time))
uber_2014$Minute <- factor(minute(uber_2014$Date.Time))
uber_2014$Second <- factor(second(uber_2014$Date.Time))
datatable(head(uber_2014))
```

##Uber Trips made Every Day in New York City
```{r, message=FALSE}
library(dplyr)
library(DT)
library(ggplot2)
library(RColorBrewer)
cols <- brewer.pal(8, "Paired")
pal <- colorRampPalette(cols)
trips_day <- group_by(uber_2014, Day) %>% summarise(Trips = n())
datatable(trips_day)
ggplot(trips_day, aes(Day, Trips)) + geom_bar(stat = "identity", fill = "maroon") +
    ggtitle("No. of Uber Trips by Day made in New York City")
trips_day_month <- group_by(uber_2014, Day, Month) %>% summarise(Trips = n())
ggplot(trips_day_month, aes(Day, Trips, fill = Month)) + geom_bar(stat = "identity") +
    ggtitle("No. of Uber Trips by Day & Month made in New York City") + scale_y_continuous() +
      scale_fill_manual(values = pal(6)) + facet_grid(Month ~ .)
```

##Uber Trips made Every Hour in New York City

```{r, message=FALSE}
library(dplyr)
library(DT)
library(ggplot2)
library(RColorBrewer)
cols <- brewer.pal(8, "Dark2")
pal <- colorRampPalette(cols)
trips_hour <- group_by(uber_2014, Hour) %>% summarise(Trips = n())
datatable(trips_hour)
ggplot(trips_hour, aes(Hour, Trips)) + geom_bar(stat = "identity", fill = "orange") + ggtitle("No. of Uber Trips Happening in New York City Every Hour")
trips_hour_month <- group_by(uber_2014, Hour, Month) %>% summarise(Trips = n())
ggplot(trips_hour_month, aes(Hour, Trips, fill = Month)) + geom_bar(stat = "identity") + ggtitle("No. of Uber Trips Happening in New York City Every Hour for Different Months") + scale_y_continuous() + scale_fill_manual(values = pal(6)) + facet_grid(Month ~ .)
```

##Uber Trips made According to The Days in New York City

```{r, fig.width=8}
cols <- brewer.pal(8, "Spectral")
pal <- colorRampPalette(cols)
trips_wday <- group_by(uber_2014, Weekday) %>% summarise(Trips = n())
datatable(trips_wday)
ggplot(trips_wday, aes(Weekday, Trips)) + geom_bar(stat = "identity", fill = "Darkred") + ggtitle("No. of Uber Trips Happening in New York City For Different Days of the Week")
trips_wday_month <- group_by(uber_2014, Weekday, Month) %>% summarise(Trips = n())
ggplot(trips_wday_month, aes(Weekday, Trips, fill = Month)) + geom_bar(stat = "identity", position = "dodge") + ggtitle("No. of Uber Trips Happening in New York City for WeekDays of Different Months") + scale_y_continuous() + scale_fill_manual(values = pal(6))
```

##Uber Trips According to Different BASE in New York City

```{r}
cols <- brewer.pal(6, "Accent")
pal <- colorRampPalette(cols)
trips_base <- group_by(uber_2014, Base) %>% summarise(Trips = n())
datatable(trips_base)
ggplot(trips_base, aes(Base, Trips)) + geom_bar(stat = "identity", fill = "Seagreen") + ggtitle("No. of Uber Trips for Varios Base in New York City")
trips_base_month <- group_by(uber_2014, Base, Month) %>% summarise(Trips = n())
ggplot(trips_base_month, aes(Base, Trips, fill = Month)) + geom_bar(stat = "identity", position = "dodge") + ggtitle("No. of Uber Trips for Diferent Base-Month in New York City") + scale_y_continuous() + scale_fill_manual(values = pal(6))
```

##Heatmap for Uber Trips by Day & Hour

```{r}
trips_hour_day <- group_by(uber_2014, Hour, Day) %>% summarise(Trips = n())
ggplot(trips_hour_day, aes(Day, Hour, fill = Trips)) + ggtitle("Heatmap of Uber Trips by Day and Hour") + geom_tile(color = "white")
```

##Heatmap for Uber Trips by Day & Month

```{r}
##library(reshape2)
##library(scales)
##library(plyr)
##cols <- brewer.pal(4, "Oranges")
##pals <- colorRampPalette(cols)
##trips_day_monthm <- melt(trips_day_month)
##trips_day_monthm <- ddply(trips_day_monthm, .(Day), transform, rescale = rescale(value))
ggplot(trips_day_month, aes(Day, Month, fill = Trips)) + ggtitle("Heatmap of Uber Trips by Days of Months") + geom_tile(color = "white")
##+ scale_fill_gradient(guide = "colourbar", low = "gray87", high = "gray35") + theme(legend.position = ##"none")
```

##Heatmap for Uber Trips by BASE and Day

```{r}
trips_base_wday <- group_by(uber_2014, Base, Weekday) %>% summarise(Trips = n())
ggplot(trips_base_wday, aes(Base, Weekday, fill = Trips)) + ggtitle("Heatmap of Uber Trips by BASE and Day") + geom_tile(color = "white")
```

##Heatmap for Uber Trips by Base and Month

```{r}
ggplot(trips_base_month, aes(Base, Month, fill = Trips)) + ggtitle("Heatmap of Uber Trips by BASE and Month") + geom_tile(color = "white")
```

##*Synopsis*

***Drawing the Inference from the above Analysis, we can state that***

1. No. of Uber Trips increase as the Days progress in a month.

2. The Month of **September, 2014** has seen the **most** Rides.

3. The Time between **15hrs to 19hrs** has the **most** Rides. In other words the Trips increase in this Time frame.

4. **Thursdays** and **Fridays** observe the **highest** amount of rides as compared to other days.

5. **B02598** and **B02617** bases have the **most** amout of **rides** during Apr'14 to Sep'14.

6. For **B02512** the no. of Rides have almost remained **constant**.

7. *September* was the *best* month for **B02764** & **B02617**.

8. *May* was the *best* month for **B02598**.

9. *April* was the *best* month for **B02682**.

10. The time between **0hrs to 4hrs** see the **least** number of rides.

##Mapping the Uber Trips

```{r, warning=FALSE}
library(ggthemes) 
min_lat <- 40.5774
max_lat <- 40.9176
min_long <- -74.15
max_long <- -73.7004

ggplot(uber_2014, aes(x=Lon, y=Lat, color = Base)) + geom_point(size = 0.06) + scale_x_continuous(limits=c(min_long, max_long)) +
scale_y_continuous(limits=c(min_lat, max_lat)) + theme_map()
```